Neutral ISM, Lyman-Alpha and Lyman-continuum in nearby starburst Haro 11 \footnote{Based on observations with HST-COS, program GO 13017, PI Heckman}

### 2017-01-16 Paper/Paper.tex
submitted version

\begin{abstract}
Star forming galaxies are believed to be a major source of ionizing
radiation responsible for reionizing the early Universe. Direct
observations of escaping ionizing radiation have however been few and
with escape fractions far below what would be necessary to account for
the ionization history of the Universe. In the local Universe, only a
few handfuls of emitters have been observed with typical escape
fractions of a few percent. It seems the mechanisms regulating this
escape need to be strongly evolving with redshift, if galaxies are to be
held responsible for reionization. Gas mass, star formation feedback and
thus star formation activity are among the main suspect parameters,
known to both regulate neutral gas coverage and evolve with cosmic time.
In this paper, we present a reanalysis of far-UV HST-COS spectra of the
first detected Lyman-continuum leaker in the local Universe, Haro 11, in
which we examine the connections between Lyman-continuum leakage and
Lyman-$\alpha$ line shape and feedback-influenced neutral ISM properties
like kinematics, geometry and column density. We discuss the two
extremes of an optically thin, density bounded ISM and a riddled,
optically thick, ionization bounded ISM and how Haro 11 fits into the
theoretical predictions made from these models. We find that the most
likely ISM model for Haro 11 knot C is one of clumpy neutral medium
embedded in a highly ionized inter-clump medium with a combined covering
fraction of unity and a residual neutral gas column density in the
ionized medium of
$5 \times 10^{13} \leq N_{\rm HI} \leq 4 \times 10^{15}$; high enough to
be optically thick to Lyman-$\alpha$, but low enough to be at least
partly transparent to Lyman-continuum and undetected in Si\textsc{ii}.
\end{abstract}

@@

Even though star forming galaxies are believed to give rise to the
majority of radiation reionizing the Universe at early times, it remains
unknown which physical conditions can facilitate the escape of the
necessary amounts of radiation from these galaxies, which contain large
amounts of neutral gas opaque to this radiation at higher column
densities. Searches for leaking galaxies at redshifts $z \gtrsim 1$ have
yielded few detections
\citep[e.g.][]{Iwata2009, Vanzella2010, Vanzella2012, Nestor2013, Cowie2009, Siana2010},
at escape fractions well below the $\sim 20\%$ needed to account for
cosmic reionization \citep{Bouwens2011, Robertson2013}. A population of
lower mass and lower luminosity, star forming galaxies are expected to
contribute to reionization, too, but high star formation is usually
coincident with higher neutral gas mass, which would cause a higher
probability of blocking the ionizing photons
\citep[e.g.][]{Erb2016, Robertson2013}. In the local Universe, only 9
leakers have been detected so far
\citetext{\citealp{Bergvall2006}; \citealp{Leitet2011}; \citealp{Leitet2013}; \citealp{Borthakur2014}; \citealp[Izotov2016Nat;][]{Izotov2016MNRAS}; \citealp{Leitherer2016}},
at escape fractions ranging typically between 1-8\%, with one as high as
$f_{\mathrm{esc}} \approx 13\%$ \citep{Izotov2016MNRAS}.

Models of ISM surrounding a central source and allowing escape of Lyman
continuum (LyC), span the range of two extremes: In one regime, the
surrounding gas is optically thin, highly ionized and density-bounded
\citep[see e.g.][]{Jaskot2013}, allowing unblocked escape for at least a
fraction of ionizing photons. In the other regime, the central source is
surrounded by an optically thick, ionization bounded medium with the
neutral medium surrounding the central Strömgren sphere not completely
covering all lines of sight to the background source. This is what is
called the \emph{picket fence model} by \citet{Heckman2011}, or the
\emph{riddled ionization bounded medium} by \citet{Verhamme2015}. The
latter paper presents modeling of the imprints of these two extreme
scenarios on the observed spectral signature of Lyman-$\alpha$ emission
lines, and suggest how these can help point to candidate LyC leakers.
The authors compare their theoretical predictions to sample of
Lyman-$\alpha$ profiles, including a section of the spectrum treated in
this work.

The present HST-COS observation is part of the sample in
\citet{Heckman2015} and \citet{Alexandroff2015}, in which details about
observation and data reduction are described in depth; we point the
reader there for further information about these. The Ly$\alpha$ profile
in this spectrum is discussed in \citet{JaskotOey}.

@@

Recently, the galaxy has been observed in 21 cm H\textsc{i} emission
with the 100 m. Robert C. Byrd Green Bank Telescope
\citep{Pardy2016arXiv}. The authors find that Haro 11 has a low gas mass
and a very low neutral gas to stellar mass fraction, and that it
contains around twice as much ionized as neutral Hydrogen.
Interestingly, the authors also find that the bulk of the neutral gas is
redshifted wrt. the systemic velocity defined from nebular emission from
the H\textsc{ii} regions surrounding the main starbursts, signifying
that the majority of the neutral gas reserves are detached from the star
formation activity, possibly a tidal arm being flung outwards as a part
of the ongoing merger event.

In this paper, we re-analyze the HST-COS spectrum of Haro 11 C acquired
as part of HST program GO 13017, PI Timothy Heckman. It was first
published in \citet{Alexandroff2015} as part of a sample of 22 Lyman
Break Analog (LBA) galaxies analyzed individually and as a stack. These
authors mainly focus on three indirect indicators of LyC leakage
\citep{Overzier2009, Heckman2011}: Residual flux in the trough of
saturated ISM absorption lines, blueshifted emission in Lyman-$\alpha$,
and weak optical {[}\ion{S}{2}{]} emission lines.

@@

To estimate the effective resolution for this target, we proceeded as
follows. We used archival imaging data of Haro 11 in the UV-continuum
from \citet{Ostlin2009}; \citet{Hayes2009}; since the absorption lines
treated here have resolution determined by the properties of their
surrounding continuum. We the extracted a circular image at the same
position and radius as the COS aperture, which was rotated to match the
orientation of the COS as given in the spectrum headers. We then
modelled vignetting in the COS aperture by multiplying each pixel in the
circular UV-continuum image with an interpolation of the values given in
the throughput grid in Fig. 7 in \citet{CosImaging}. Finally, the
aperture cutouts were collapsed along the cross-dispersion direction,
and the FWHM of the flux distribution along the dispersion direction is
reported as the effective seeing of this observation. This is converted
from arc seconds to COS resolution elements by the factor 0.171
arcsec/resel reported in the COS instrument handbook \citep{CosHandbook}

@@

Figure~\ref{fig:SingleLines} shows the individual profiles of the
transitions included in our analysis; the upper panel shows transitions
of \ion{Si}{2}, lower panel of \ion{Si}{4}. It is plainly visible that
the ionization fraction is high, with the \ion{Si}{4} curves being
considerably deeper than the low-ionized lines. Looking at the upper
panel, \ion{Si}{2} $\lambda \lambda 1304, 1526$ are somewhat shallower
than \ion{Si}{2} $\lambda 1260$. The former two lines have comparable
oscillator strengths, both about a factor of 10 lower than that of
$\lambda 1260$. It is thus clear that we do not find ourselves in the
optically thin regime, in which the latter line should be
correspondingly around 10 times stronger; on the other hand, it is
possible that the two weak lines are not completely saturated. In the
lower panel, the two \ion{Si}{4} lines have oscillator strengths within
a factor of 2 of each other. They are thus at first glance consistent
with a medium that is not completely opaque, but not with an optically
thin one, and within uncertainties consistent with the optically thick.
The stronger absorption in \ion{Si}{4} reveals a high level of
ionization of the medium covering the central cluster.

@@

In any given velocity bin, the line intensity in terms of the continuum
intensity is given as
%
\begin{equation}
\label{eq:II0}
\frac{I}{I_0} = 1 - f_C (1 - e^{-\tau}),
\end{equation}
%
 with the optical depth $\tau$ given as:
%
\begin{equation}
\label{eq:tau}
\tau = f\lambda \frac{\pi e^2}{m_e c} N 
       = f\lambda \frac{N}{3.768 \times 10^{14}}
\end{equation}
%
 Here, $f$ is the oscillator strength of a given transition, $\lambda$
is its rest frame wavelength in Å, $N = N(v)$ is the column density of
the relevant ion \emph{within the given velocity bin}, and $f_{C}$ is
the covering fraction of neutral gas \emph{moving at the given
velocity}. When multiple absorption lines are present which arise from
the same ground state; the population of this state is the same for all
transitions, and their relative strengths are governed simply by their
oscillator strengths, and with two or more such transitions, $f_{C}$ and
$N$ can be inferred from knowledge of $f \lambda$ and measured values of
$I/I_0$.

@@

The picture is what we would expect from a Lyman Continuum leaker: weak
neutral and strong ionized absorption profiles, the latter almost
everywhere consistent with full coverage, reveal a highly ionized medium
in front of the central cluster. This is also in good agreement with the
findings of \citet{Pardy2016arXiv}, who conclude that the galaxy has
about twice as much ionized as neutral gas. In addition to this, it is
interesting to note the close similarity in shape between the
\ion{Si}{2} and \ion{Si}{4} line profiles, indicating that they likely
represent two different phases in the same higher density regions. These
regions will likely be ionized on the side facing the central cluster,
being photoionized by its hot, massive stars. This also implies that the
nebular emission outflow found in {[}Rivera-Thorsen et al, submitted{]}
at least partially traces the same gas as the \ion{Si}{4} absorption
here and, by extension, also the neutral medium. Looking at
fig.~\ref{fig:SingleLines}, we see that maximum absorption and thus the
largest amount of gas in a single velocity bin for both phases is found
at around $v \approx -50$ km s⁻¹, where also a component is found in
{[}Rivera-Thorsen et al., submitted{]}. The outflow velocity
$v_{\text{int}}$ is at 149 ± 7 km s⁻¹, fully consistent with the
findings of \citet{Heckman2015} and \citet{Alexandroff2015}.
Interestingly, \citet{Sandberg2013} found from the neutral Sodium
resonance absorption doublet $\lambda \lambda 5889.95,5895.92$ (NaD) an
overall, weak \emph{redshift} of $v = 32$ km s⁻¹. From
e.g.~fig.~\ref{fig:HisLisLya}, it is evident that while neutral gas
\emph{is} present at these velocities, the velocity is at odds with the
integrated velocity of $v = -149$ km s⁻¹ found in this work. We note
that Na \textsc{i} has a very low ionization potential of only $\sim 5$
eV, meaning that these atoms may well only be present in the densest
and/or dustiest regions, in which Sodium is shielded from ionization.
This is in agreement with the finding of \citet{Sandberg2013} that NaD,
despite being a strong transition, shows absorption of only $\sim 95\%$
of continuum level and is mostly found in small, optically thick clouds.
This suggests that the NaD is tracing only the densest and/or dustiest
regions which are more slowly accelerated by star formation feedback
than the surrounding, more dilute medium. In fig.~\ref{fig:apert}, dusty
regions in the aperture are apparent E and W of knot C, and in the
better resolution of fig. 1 in \citet{Adamo2010}, these regions seem
like they might be connected by a narrow dust lane partially covering
the background source. We speculate that the NaD absorption of
\citet{Sandberg2013} might be associated with this.

Also the absorption feature in the Lyman $\alpha$ profile in the upper
panel seems to morphologically follow the shape of the metal lines,
indicating that radiative transfer effects are modest, indicative of a
fairly low column density of HI around line center. We find a Lyman
$\alpha$ peak velocity of $v_{\rm peak}^{\mathrm{Ly}\alpha} = 158 \pm 1$
km s⁻¹ w.r.t the systemic velocity found by \citet{Sandberg2013},
derived from nebular emission lines in the region around knot C. This
velocity, as it is also discussed in \citet{Verhamme2015}, is just
consistent with their theoretical predictions for a density-bounded,
low-column density system, albeit on the upper limit of their allowed
range.

@@

Lyman $\alpha$ escape is mainly governed by gas at or near systemic
velocity. In the middle panel of fig.~\ref{fig:WithColDens} (a) is shown
the best-fit column density $N_{\rm Si II}$ for each velocity bin. The
value at systemic velocity is $\log_{10}(N_{\rm Si II}) = 12.1 \pm 0.2$.
We adapt the value for local starbursts of log(Si/O) $= -1.59 \pm 0.07$
from \citet{Garnett1995}, and use this to estimate the column density of
neutral Hydrogen in front of the light source in the same way is in
{[}Puschnig et al., submitted to MNRAS{]} as follows. \citet{Guseva2012}
found a metallicity of Haro 11 C of $12 + \log_{10}(\text{O/H}) = 8.1$.
With $\log_{10}$ Si/O $= -1.59 \pm 0.07$, this leads to a Si/H ratio in
the neutral medium of Si / H $= 3.24^{+0.57}_{-0.48} 10^{-6}$, leading
to a Hydrogen column density of $6.2^{+0.9}_{-1.1} \times 10^{17}$
cm\textsuperscript{-2}. Since HI gets opaque to ionizing radiation at
$\log N \sim 17.2$ \citep{Verhamme2015}, this range is not consistent
with the low- optical depth, density bounded scenario. Furthermore;
while Ly$\alpha$ radiative transfer is dominated by gas of $v \sim v_0$,
Lyman continuum is sensitive to \ion{H}{1} at \emph{all} velocities.
Integrating the computed column densities over $-450$ km s⁻¹ $< v < 150$
km s⁻¹, the velocity range over which the column densities can be
reasonably well determined (and removing the unphysically high value in
the bin at $v \approx 222$ km s⁻¹), yields a conservative estimate of
$\log N_{\rm Si II} \approx 19.4$, corresponding to more than 150
optical depths in Lyman Continuum, strongly incompatible with an
optically thin, density-bounded scenario.

However, the pure riddled ionization bounded scenario is easily ruled
out since the Ly$\alpha$ profile does not have any appreciable emission
component at zero velocity. We therefore expect a residual neutral
fraction to remain in the ionized phase; a fraction which has a column
density high enough to block Ly$\alpha$ efficiently at line center, but
low enough to be at least part transparent to Lyman Continuum radiation
and undetected in \ion{Si}{2}.

@@

The existence of a diffuse neutral component being present in the
ionized medium seems consistent with what is found in the Green Pea
galaxies of \citep{Henry2015} - these galaxies have sometimes very low
LIS optical depth, which is usually indicative of low covering fractions
and high porosity of the neytral medium - and yet they find covering
fractions of near unity of \ion{H}{1} from Lyman $\beta$ absorption,
indicating column densities of $N \gtrsim 10^{16}$. Like in the case of
Haro 11, this indicates that a non-negligible neutral component must be
present in the ionized phase, although it may very well be of such low
densities and metallicity that metal absorption from this medium is
undetectable.

@@

The exact value of the metallicity is however uncertain. The values
found from nebular recombination lines by \citet{Guseva2012} are
measured mainly in the central HII regions around the clusters; and
differ by 0.2 dex between knot B and C. The neutral, outflowing gas
could be mixed, or have an unseen LOS distance component larger than the
knot separation, drawing into question which is the better value to
assume for this gas. We base our conclusions on the value found for knot
C, but note that using the metallicity of $12 + \log (O/H) = 8.3$ found
for knot B by \citet{Guseva2012}, the inferred H \textsc{i} column
densities are $3.9 \pm 0.6 \times 10^{17}$ cm\textsuperscript{-2}.
However, the question of the exact metallicity of the outflowing gas is
complicated. Gas closer to the star forming regions is expected to be
more strongly enriched than gas further away, which would imply that the
HI column density is \emph{larger} than inferred from \ion{Si}{2} above.
To this can be added the further complication stemming from the merger
event that the galaxy is currently undergoing, which may have mixed gas
of different metal contents. In any case, though, we would expect the
regions nearest to the starbursts to generally have higher metallicity
than the surrounding cool gas, such that the HI column density inferred
above is more likely to be underestimated than overestimated.

@@

The conclusions drawn by the AOD method rest on the assumption that the
observed lines are close to being pure absorption lines, with
redistribution of photons due to radiative transfer effects being
modest. Modelling work by e.g. \citet{Prochaska2011} and
\citet{Scarlata2015} has shown that in certain conditions, radiative
transfer can re-fill absorption features in a way that can make an
isotropic, optically thin mimic the observational fingerprints of a
system of optically thick clumps. We therefore need to investigate
whether our conclusions could be generated by such effects.

@@

Because the ground level in \ion{Si}{2} is a doublet with a short-lived
upper level, each of the Silicon lines included in the analysis has a
fluorescent emission line with practically no absorption component. The
fluorescent line can be used to constrain the possible effect of
radiative infilling of the absorption trough. If each photon in the
fluorescent transition on average has scattered once in the neutral
medium, the ratio of emission in this line and re-filling of the
resonant absorption line is simply that of their Einstein coefficients
$A_{ki}$. For each subsequent scattering, more photons escape directly
through the fluorescent channel, and the infilling of the absorption
line further suppressed.

In fig.~\ref{fig:rt}, the observed line profile is shown in blue, along
with the best fit to \ion{Si}{2}$^*$ shown in orange. Since the
fluorescent and resonant reemission originate in the same environments,
the shape of the resonant reemission profile can be completely
determined from the shape and center of the fluorescent line and the
assumption of one scattering. It is therefore simple to construct this
theoretical reemission line, subtract it from the observed absorption
line and find a lower limit for the depth of the pure absorption line
without reemission; this model line sis shown in gray in the figure. It
is readily seen that the difference between theoretical and observed
line is so small that it cannot affect our conclusions about the
covering factors and column densities of the neutral gas significantly.

We have not performed similar measurements for the lines at
$\lambda \lambda 1304, 1526$, but note that if significant refilling is
present here, this would mean the column densities had actually been
underestimated in our analysis, which would strengthen the conclusions
of a clumpy medium optically thick in \ion{Si}{2} $1260$, but part
transparent in the much weaker other lines.

@@

Whether this is a physically reasonable strength is difficult to say
with precision, but we note that to reproduce the feature observed,
$\lambda 1334$ must be optically thick and $\lambda 1335$ part
transparent. This leave open a degeneracy of relative level populations
and ion abundance, such that the choice of this value only weakly
constrains these quantities and thus it is compatible with a wide range
of scenarios.

>>>>

\begin{abstract}
Star forming galaxies are believed to be a major source of ionizing
radiation responsible for reionizing the early Universe. Direct
observations of escaping ionizing radiation have however been few and
with escape fractions far below what would be necessary to account for
the ionization history of the Universe. In the local Universe, only a
few handfuls of emitters have been observed with typical escape
fractions of a few percent. It seems the mechanisms regulating this
escape need to be strongly evolving with redshift, if galaxies are to be
held responsible for reionization. Gas mass, star formation feedback and
thus star formation activity are among the main suspect parameters,
known to both regulate neutral gas coverage and evolve with cosmic time.
In this paper, we present a reanalysis of far-UV HST-COS spectra of the
first detected Lyman-continuum leaker in the local Universe, Haro 11, in
which we examine the connections between Lyman-continuum leakage and
Lyman-$\alpha$ line shape and feedback-influenced neutral ISM properties
like kinematics, geometry and column density. We discuss the two
extremes of an optically thin, density bounded ISM and a riddled,
optically thick, ionization bounded ISM and how Haro 11 fits into the
theoretical predictions made from these models. We find that the most
likely ISM model for Haro 11 knot C is one of clumpy neutral medium
embedded in a highly ionized inter-clump medium with a combined covering
fraction of unity and a residual neutral gas column density in the
ionized medium of
$5 \times 10^{13} \leq N_{\rm HI} \leq 4 \times 10^{15}$ cm⁻²; high
enough to be optically thick to Lyman-$\alpha$, but low enough to be at
least partly transparent to Lyman-continuum and undetected in
\ion{Si}{2}.
\end{abstract}

@@

Even though star forming galaxies are believed to give rise to the
majority of radiation reionizing the Universe at early times, it remains
unknown which physical conditions can facilitate the escape of the
necessary amounts of radiation from these galaxies, which contain large
amounts of neutral gas opaque to this radiation at higher column
densities. Searches for leaking galaxies at redshifts $z \gtrsim 1$ have
yielded few detections
\citep[e.g.][]{Iwata2009, Vanzella2010, Vanzella2012, Nestor2013, Cowie2009, Siana2010},
at escape fractions well below the $\sim 20\%$ needed to account for
cosmic reionization \citep{Bouwens2011, Robertson2013}. A population of
lower mass and lower luminosity, star forming galaxies are expected to
contribute to reionization, too, but high star formation is usually
coincident with higher neutral gas mass, which would cause a higher
probability of blocking the ionizing photons
\citep[e.g.][]{Erb2016, Robertson2013}. In the local Universe, only 9
leakers have been detected so far
\citep{Bergvall2006, Leitet2011, Leitet2013, Borthakur2014, Izotov2016Nat, Izotov2016MNRAS, Leitherer2016},
at escape fractions ranging typically between 1-8\%, with one as high as
$f_{\mathrm{esc}} \approx 13\%$ \citep{Izotov2016MNRAS}.

Models of ISM surrounding a central source and allowing escape of Lyman
continuum (LyC), span the range of two extremes: In one regime, the
surrounding gas is optically thin, highly ionized and density-bounded
\citep[see e.g.][]{Jaskot2013}, allowing unblocked escape for at least a
fraction of ionizing photons. In the other regime, the central source is
surrounded by an optically thick, ionization bounded medium with the
neutral medium surrounding the central Strömgren sphere not completely
covering all lines of sight to the background source \citep[see
e.g.][]{Zackrisson2013}. This is what is called the \emph{picket fence
model} by \citet{Heckman2011}, or the \emph{riddled ionization bounded
medium} by \citet{Verhamme2015}. The latter paper presents modeling of
the imprints of these two extreme scenarios on the observed spectral
signature of Lyman-$\alpha$ emission lines, and suggest how these can
help point to candidate LyC leakers. The authors compare their
theoretical predictions to sample of Lyman-$\alpha$ profiles, including
a section of the spectrum treated in this work.

@@

Recently, the galaxy has been observed in 21 cm \ion{H}{1} emission with
the 100 m. Robert C. Byrd Green Bank Telescope \citep{Pardy2016arXiv}.
The authors find that Haro 11 has a low gas mass and a very low neutral
gas to stellar mass fraction, and that it contains around twice as much
ionized as neutral Hydrogen. Interestingly, the authors also find that
the bulk of the neutral gas is redshifted wrt. the systemic velocity
defined from nebular emission from the \ion{H}{2} regions surrounding
the main starbursts, signifying that the majority of the neutral gas
reserves are detached from the star formation activity, possibly a tidal
arm being flung outwards as a part of the ongoing merger event.

In this paper, we re-analyze the HST-COS spectrum of Haro 11 C acquired
as part of HST program GO 13017, PI Timothy Heckman. It was first
published in \citet{Alexandroff2015} and \citet{Heckman2015} as part of
a sample of 22 Lyman Break Analog (LBA) galaxies analyzed individually
and as a stack. These authors mainly focus on three indirect indicators
of LyC leakage \citep{Overzier2009, Heckman2011}: Residual flux in the
trough of saturated ISM absorption lines, blueshifted emission in
Lyman-$\alpha$, and weak optical {[}\ion{S}{2}{]} emission lines.
Details about observation and data reduction are described in depth in
\citet{Alexandroff2015}; we point the reader there for further
information about these. The Ly$\alpha$ profile in this spectrum is
discussed in \citet{Verhamme2015}.

@@

To estimate the effective resolution for this target, we proceeded as
follows. We used archival imaging data of Haro 11 in the UV-continuum
from \citet{Ostlin2009}; \citet{Hayes2009}; since the absorption lines
treated here have resolution determined by the properties of their
surrounding continuum. We then extracted a circular image at the same
position and radius as the COS aperture, which was rotated to match the
orientation of the COS as given in the spectrum headers. We modelled
vignetting in the COS aperture by multiplying each pixel in the circular
UV-continuum image with an interpolation of the values given in the
throughput grid in Fig. 7 in \citet{CosImaging}. Finally, the aperture
cutouts were collapsed along the cross-dispersion direction, and the
FWHM of the flux distribution along the dispersion direction is reported
as the effective resolution of this observation. This is converted from
arc seconds to COS resolution elements by the factor 0.171 arcsec/resel
reported in the COS instrument handbook \citep{CosHandbook}

@@

Figure~\ref{fig:SingleLines} shows the individual profiles of the
transitions included in our analysis; the upper panel shows transitions
of \ion{Si}{2}, lower panel of \ion{Si}{4}. It is plainly visible that
the ionization fraction is high, with the \ion{Si}{4} curves being
considerably deeper than the low-ionized lines. Looking at the upper
panel, \ion{Si}{2} $\lambda \lambda 1304, 1526$ are somewhat shallower
than \ion{Si}{2} $\lambda 1260$. The former two lines have comparable
oscillator strengths, both about a factor of 10 lower than that of
$\lambda 1260$. It is thus clear that we do not find ourselves in the
optically thin regime, in which the latter line should be
correspondingly around 10 times stronger; on the other hand, it is
possible that the two weak lines are not completely saturated. In the
lower panel, the two \ion{Si}{4} lines have oscillator strengths within
a factor of 2 of each other. They are thus at first glance consistent
with a medium that is not completely opaque, but not with an optically
thin one, and within uncertainties consistent with the optically thick.
The stronger absorption in \ion{Si}{4} reveals a high level of
ionization of the medium covering the background source.

@@

In any given velocity bin, the residual line intensity in terms of the
continuum intensity is given as
%
\begin{equation}
\label{eq:II0}
\frac{I}{I_0} = 1 - f_C (1 - e^{-\tau}),
\end{equation}
%
 with the optical depth $\tau$ given as:
%
\begin{equation}
\label{eq:tau}
\tau = f\lambda \frac{\pi e^2}{m_e c} N 
       = f\lambda \frac{N}{3.768 \times 10^{14}}
\end{equation}
%
 Here, $f$ is the oscillator strength of a given transition, $\lambda$
is its rest frame wavelength in Å, $N = N(v)$ is the column density of
the relevant ion \emph{within the given velocity bin}, and $f_{C}$ is
the covering fraction of neutral gas in same velocity bin. When multiple
absorption lines are present which arise from the same ground state; the
population of this state is the same for all transitions, and their
relative strengths are governed simply by their oscillator strengths,
and with two or more such transitions, $f_{C}$ and $N$ can be inferred
from knowledge of $f \lambda$ and measured values of $I/I_0$.

@@

The picture is what we would expect from a Lyman Continuum leaker: weak
neutral and strong ionized absorption profiles, the latter almost
everywhere consistent with full coverage, reveal a highly ionized medium
in front of the central cluster. This is also in good agreement with the
findings of \citet{Pardy2016arXiv}, who conclude that the galaxy has
about twice as much ionized as neutral gas. In addition to this, it is
interesting to note the close similarity in shape between the
\ion{Si}{2} and \ion{Si}{4} line profiles, indicating that they likely
represent two different phases in the same higher density regions. These
regions will likely be ionized on the side facing the central cluster,
being photoionized by its hot, massive stars. This also implies that the
nebular emission outflow found in {[}Rivera-Thorsen et al, submitted{]}
at least partially traces the same gas as the \ion{Si}{4} absorption
here and, by extension, also the neutral medium. Looking at
fig.~\ref{fig:SingleLines}, we see that maximum absorption and thus the
largest amount of gas in a single velocity bin for both phases is found
at around $v \approx -50$ km s⁻¹, where also a component is found in
{[}Rivera-Thorsen et al., submitted{]}. The outflow velocity
$v_{\text{int}}$ is at 149 ± 7 km s⁻¹, fully consistent with the
findings of \citet{Heckman2015} and \citet{Alexandroff2015}.
Interestingly, \citet{Sandberg2013} found from the neutral Sodium
resonance absorption doublet $\lambda \lambda 5889.95,5895.92$ (NaD) an
overall, weak \emph{redshift} of $v = 32$ km s⁻¹. From
e.g.~fig.~\ref{fig:HisLisLya}, it is evident that while neutral gas
\emph{is} present at these velocities, the velocity is at odds with the
integrated velocity of $v = -149$ km s⁻¹ found in this work. We note
that \ion{Na}{1} has a very low ionization potential of only $\sim 5$
eV, meaning that these atoms may well only be present in the densest
and/or dustiest regions, in which Sodium is shielded from ionization.
This is in agreement with the finding of \citet{Sandberg2013} that NaD,
despite being a strong transition, shows absorption of only $\sim 95\%$
of continuum level and is mostly found in small, optically thick clouds.
This suggests that the NaD is tracing only the densest and/or dustiest
regions which are more slowly accelerated by star formation feedback
than the surrounding, more dilute medium. In fig.~\ref{fig:apert}, dusty
regions in the aperture are apparent E and W of knot C, and in the
better resolution of fig. 1 in \citet{Adamo2010}, these regions seem
like they might be connected by a narrow dust lane partially covering
the background source. We speculate that the NaD absorption of
\citet{Sandberg2013} might be associated with this.

Also the absorption feature in the Lyman $\alpha$ profile in the upper
panel seems to morphologically follow the shape of the metal lines,
indicating that radiative transfer effects are modest, indicative of a
fairly low column density of \ion{H}{1} around line center. We find a
Lyman $\alpha$ peak velocity of
$v_{\rm peak}^{\mathrm{Ly}\alpha} = 158 \pm 1$ km s⁻¹ w.r.t the systemic
velocity found by \citet{Sandberg2013}, derived from nebular emission
lines in the region around knot C. This velocity, as it is also
discussed in \citet{Verhamme2015}, is just consistent with their
theoretical predictions for a density-bounded, low-column density
system, albeit on the upper limit of their allowed range.

@@

Lyman $\alpha$ escape is mainly governed by gas at or near systemic
velocity. In the middle panel of fig.~\ref{fig:WithColDens} (a) is shown
the best-fit column density $N_{\rm Si II}$ for each velocity bin. The
value at systemic velocity is $\log_{10}(N_{\rm Si II}) = 12.1 \pm 0.2$.
We adapt the value for local starbursts of log(Si/O) $= -1.59 \pm 0.07$
from \citet{Garnett1995}, and use this to estimate the column density of
neutral Hydrogen in front of the light source in the same way is in
{[}Puschnig et al., submitted to MNRAS{]} as follows. \citet{Guseva2012}
found a metallicity of Haro 11 C of $12 + \log_{10}(\text{O/H}) = 8.1$.
With $\log_{10}$ Si/O $= -1.59 \pm 0.07$, this leads to a Si/H ratio in
the neutral medium of Si / H $= 3.24^{+0.57}_{-0.48} 10^{-6}$, leading
to a Hydrogen column density of $6.2^{+0.9}_{-1.1} \times 10^{17}$ cm⁻²
for just one velocity bin. Since HI gets opaque to ionizing radiation at
$\log N \sim 17.2$ \citep{Verhamme2015}, this range is not consistent
with the low- optical depth, density bounded scenario.

Furthermore; while Ly$\alpha$ radiative transfer is dominated by gas of
$v \sim v_0$, Lyman continuum is sensitive to \ion{H}{1} at \emph{all}
velocities. The total column density of \ion{Si}{2} depends on the
configuration of the neutral clouds. Assuming that the area covered by
gas in each velocity bin is completely randomly located, a lower limit
to the total column density is:
%
\begin{equation}
\label{NHImean}
N_{\rm Si II}^{\rm mean} = \sum_i N_{\text{HI}, i} f_{C, i}
\end{equation}
%
 Summing this over $-450$ km s⁻¹ $< v < 150$ km s⁻¹, the velocity range
over which the column densities can be reasonably well determined (and
removing the unphysically high value in the bin at $v \approx 222$ km
s⁻¹), yields a lower limit of $\log N_{\rm Si II} = 13.52 \pm 0.15$ and
$\log N_{\rm H I} 19.01 \pm 0.17$, corresponding to $\approx 64$ optical
depths in Lyman Continuum, strongly incompatible with an optically thin,
density-bounded scenario.

However, the \emph{pure} riddled ionization bounded scenario is easily
ruled out since the Ly$\alpha$ profile does not have any appreciable
emission component at zero velocity. We therefore expect a residual
neutral fraction to remain in the ionized phase; a fraction which has a
column density high enough to block Ly$\alpha$ efficiently at line
center, but low enough to be at least part transparent to Lyman
Continuum radiation and undetected in \ion{Si}{2}.

@@

The existence of a diffuse neutral component being present in the
ionized medium seems consistent with what is found in the Green Pea
galaxies of \citep{Henry2015} - these galaxies have sometimes very low
LIS absorption depths, which is usually indicative of low covering
fractions and high porosity of the neutral medium - and yet they find
covering fractions of near unity of \ion{H}{1} from Lyman $\beta$
absorption, indicating column densities of $N \gtrsim 10^{16}$. Like in
the case of Haro 11, this indicates that a non-negligible neutral
component must be present in the ionized phase, although it may very
well be of such low densities and metallicity that metal absorption from
this medium is undetectable.

@@

The exact value of the metallicity is however uncertain. The values
found from nebular recombination lines by \citet{Guseva2012} are
measured mainly in the central \ion{H}{2} regions around the clusters;
and differ by 0.2 dex between knot B and C. The neutral, outflowing gas
could be mixed, or have an unseen LOS distance component larger than the
knot separation, drawing into question which is the better value to
assume for this gas. We base our conclusions on the value found for knot
C, but note that using the metallicity of $12 + \log (O/H) = 8.3$ found
for knot B by \citet{Guseva2012}, the inferred \ion{H}{1} column
densities are $3.9 \pm 0.6 \times 10^{17}$ cm\textsuperscript{-2}.
However, the question of the exact metallicity of the outflowing gas is
complicated. Gas closer to the star forming regions is expected to be
more strongly enriched than gas further away, which would imply that the
\ion{H}{1} column density is \emph{larger} than inferred from
\ion{Si}{2} above. To this can be added the further complication
stemming from the merger event that the galaxy is currently undergoing,
which may have mixed gas of different metal contents. In any case,
though, we would expect the regions nearest to the starbursts to
generally have higher metallicity than the surrounding cool gas, such
that the HI column density inferred above is more likely to be
underestimated than overestimated.

@@

The conclusions drawn by the AOD method rest on the assumption that the
observed lines are close to being pure absorption lines, with
redistribution of photons due to radiative transfer effects being
modest. Modelling work by e.g. \citet{Prochaska2011} and
\citet{Scarlata2015} has shown that in certain conditions, radiative
transfer can re-fill absorption features in a way that can make an
isotropic, optically thin medium mimic the observational fingerprints of
a system of optically thick clumps. We therefore need to investigate
whether our conclusions could be generated by such effects.

@@

Because the ground level in \ion{Si}{2} is a doublet with a short-lived
upper level, each of the Silicon lines included in the analysis has a
fluorescent emission companion with practically no absorption component.
The fluorescent line can be used to constrain the possible effect of
radiative infilling of the absorption trough. If each photon in the
fluorescent transition on average has scattered once in the neutral
medium, the ratio of emission in this line and re-filling of the
resonant absorption line is simply that of their Einstein coefficients
$A_{ki}$. For each subsequent scattering, more photons escape directly
through the fluorescent channel, and the infilling of the absorption
line is further suppressed.

In fig.~\ref{fig:rt}, the observed line profile is shown in blue, along
with the best fit to \ion{Si}{2}$^*$ shown in orange. Since the
fluorescent and resonant reemission originate in the same environments,
the shape of the resonant reemission profile can be completely
determined from the shape and center of the fluorescent line and the
assumption of one scattering. It is therefore simple to construct this
theoretical reemission line, subtract it from the observed absorption
line and find the limiting depth of the pure absorption line without
reemission; this modeled line is shown in gray in the figure. It is
readily seen that the difference between worst-case intrinsic and
observed line is so small that it cannot affect our conclusions about
the covering factors and column densities of the neutral gas
significantly.

We have not performed similar measurements for the lines at
$\lambda \lambda 1304, 1526$, but note that if significant refilling is
present here, this would mean the column densities had been
underestimated in our analysis, which would strengthen the conclusions
of a clumpy medium optically thick in \ion{Si}{2} $1260$, but part
transparent in the much weaker other lines.

@@

Whether this is a physically reasonable strength is difficult to say
with precision, but we note that to reproduce the feature observed,
$\lambda 1334$ must be optically thick and $\lambda 1335$ part
transparent. This leaves open a degeneracy of relative level populations
and ion abundance, such that the choice of this value only weakly
constrains these quantities and thus it is compatible with a wide range
of scenarios.
